Water discharge has a variable impact on sediment transport capacity in subglacial systems

### 2023-07-13 Delaney2023.tex
small grammer bits

\item Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more than subaerial systems.

\item The systems' divergence in transport response may impact sediment dynamics at glacier margins and interpretation of sediment records.
\end{itemize}

@@

Recheck results figure names.

@@

In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that 
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w_c\, v^2,
  \end{equation}
\end{linenomath*}
% 
where $w_c$ is the width of the channel.
Following the conservation of mass, the velocity of the water flowing through a  channel is given as 
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q_w}{S},
  \end{equation}
\end{linenomath*}
% 
where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.

@@

As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \citep{alley1997,swift2021}.
The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.
Furthermore, thorough observations of water velocity and channel width in subglacial systems are limited and not contrasted with subaerial conditions at the glacier margins.

@@

\begin{table}[ht]
  \centering
  \caption{Variables, parameters, and constants }
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
    Width-integrated shear-stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

@@

Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
The formulation here does not consider englacial storage of water, and thus that changes in hydraulic head at the top of the glacier are negligable in the evolution of the channel size. 
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_2 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n.
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the second term represents the creep closure of the channel from ice deformation.

@@

With knowledge of cross sectional area $S$, the shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
\begin{equation}
  \label{eq:tau}
  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
\end{equation}
% 
where  the water velocity $v_g = \frac{Q_w}{S_g}$.

@@

The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
Water velocity, shear-stress and width-integrated shear-stress in the subglacial and subaerial systems are presented below.

To characterize the variability in sediment discharge capacity in subglacial compared to subaerial systems further, the parameterizations are applied to range of different channel geometry factors ($\beta$ and $k$) and friction factors $f_r$ and $f_p$.
Model runs with parameter combinations  are culled if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some dye tracing measurements \citep[e.g.]{werder2010b,chandler2013}.
The standard deviation of water velocity, shear-stress and width-integrated shear-stress across the season long hydrographs are presented below.

@@

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig2.pdf}
    \caption{Model outputs as in response to hydrograph (a and e) from \alpine{} (a-d) and \icesheet{} (e-h). Gray line in a and e represent subglacial channel cross sectional area. Below this black lines represent model outputs of velocity, shear-stress and width-integrated shear-stress from the subglacial system, while grey lines represent the subaerial system.
      Plots are shown at $15$\unit{min} interval.
    Arrows denote some peak events where peak the variable peaks in the subglacial system prior to the subaerial one.} 
    \label{fig:model_outs}
  \end{figure}
\end{center}

Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear-stress and total shear-stress exhibit different seasonal evolutions and peaks of water velocity and shear stress (representing the potential for sediment mobilization) and total shear stress (representing the total sediment transport capacity; Figure~\ref{fig:model_outs}).

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig3.png}
    \caption{Relationship between water discharge and velocity, shear stress and width-integrated shear stress for the \alpine{} hydrograph (a-c) and \icesheet (d-f). Variables have been normalized to by the mean values for better comparison amongst the model runs.
      Plots are shown at $15$\unit{min} interval.} 
    \label{fig:Qw_vari}
  \end{figure}
\end{center}

For each water discharge value in the subaerial system, there is a single water velocity, shear-stress and total shear-stress (Figure~\ref{fig:Qw_vari}).
Conversely in the subglacial system,  water velocity, shear-stress and total shear-stress vary substantially for a given water discharge (Figure~\ref{fig:Qw_vari}).
For instance, over $15$ minute time periods at \alpine{}, the range of water velocities and shear-stress spans from those below the mean to those almost $2$ times the mean velocity.
Substantially high values of water velocity and shear-stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
At \icesheet, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
However, mean values of water velocity can occur at water discharges between roughly $50$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $225$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the observed range of water discharge.

When accounting for the evolving channel width of the subglacial conduit, total shear-stress generally increases with water discharge.
Yet the total shear-stress across the can very substantially, with the highest values in total shear-stress occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
The variance is less at the \icesheet case, yet a range of total shear-stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).

Over the course of the season, large peaks in water velocity and shear stress, along with total shear stress, occur in response to increases in water discharge in both systems.
In the subglacial system, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
Conversely, in the subaerial system peaks in these variables occur when water discharge is at its highest value.
As a result, peaks in the subglacial variables occur prior to the peaks in the subaerial variables, representing lags sediment mobilization and  transport capacity when sediment is transferred from on system to the other.

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig4.png}
    \caption{Variance (standard deviation) in water velocity, shear stress and total shear stress against aggregation time for the subglacial (blue) and subaerial (orange) systems from the \alpine{} (a-c) and \icesheet{} (d-f).
      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations for each the subglacial and subaerial systems.
      Solid lines denote the mean values.
      Time intervals show in markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, $15$\,\unit{d}, and $30$\,\unit{d} aggregation periods.
    } 
    \label{fig:multi_run}
  \end{figure}
\end{center}

Variability in water velocity and shear stress remain higher in the subglacial system compared to the subaerial one, across the range of parameter values examined for both systems.
However, in the \alpine{} system accounting for channel width makes the variability in the systems comparable (Figure~\ref{fig:multi_run}).
Yet, the variability over time periods less than one half day, the variability in the subglacial system is somewhat higher. 
Conversely, in \icesheet{}, with lower water discharge variability, the variability in total shear stress remains higher in the subglacial system across the time aggregation periods.

In all cases, the variability in velocity, shear stress and total shear stress decreases after $1-5$ day periods (Figure~\ref{fig:multi_run}).
This suggests that the effects of the variable response to hydrologic forcing in these cases is reduced after these time periods.
Yet, substantial differences persist in periods of up to $30$ days, as the variability might not converge given seasonal variations in water discharge.
Note that lags and time of peak sediment transport conditions will remain offset (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters is reduced with increasing time aggregation (Figure~\ref{fig:multi_run}).

@@

Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; e.g. ]{yager2018}.
In the subaerial parameterization, this means that sediment is transported across all channel widths, as opposed to large channel widths during floods when much sediment transport occurs \citep{wolman1960}.
As a result, the  the subaerial systems variability in sediment transport capacity across the range of discharge values could be underestimated in the parameterization here (Figure~\ref{fig:Qw_vari}).
Furthermore, the similar variance of total shear stress in some cases shows (Figure~\ref{fig:multi_run}) that subglacial channels contain self-organizing feedbacks that limit sediment transport capacity, as in fluvial systems \citep{phillips2016}.

In natural systems, variable mobilization and deposition of sediment may accumulate as water and sediment accumulates through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the non-dimensional parameterization.
Such variability could be especially strong in the subglacial system \citep{delaney2023}.

@@

Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronouced on ice sheets with large catchments and reduced variability.
Furthermore, total shear stress can remain high across a range of water discharges (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
In fact, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}.
Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds to the size of the glacier conduit, in addition to water discharge (Figure~\ref{fig:model_outs} and Section~\ref{sect:sub_mode}).

This suggests that larger climatic perturbations that isolate  water discharge might be needed to isolate significant changes sediment discharge from glacial systems compared to subaerial on, as the noise resulting from increased variability in sediment transport capacity must be overcome.
Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.

>>>>

\item Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more than in subaerial systems.

\item The systems' divergence in transport response may impact sediment dynamics at glacier margins and the interpretation of sediment records.
\end{itemize}

@@

%Recheck results figure names.

@@

In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that the water could carry. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that 
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w_c\, v^2,
  \end{equation}
\end{linenomath*}
% 
where $w_c$ is the width of the channel.
Following the conservation of mass, the velocity of the water flowing through a  channel is given as 
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q_w}{S},
  \end{equation}
\end{linenomath*}
% 
where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.

@@

As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
However, some suggest that subglacial sediment discharge  $Q_s \propto Q_w^{\frac{9}{2}}$, assuming that water discharge, changing channel width, and depth are each proportionally responsible for the variance of water velocity and sediment transport capacity \citep{alley1997,swift2021}.
The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.
Furthermore, thorough observations of water velocity and channel width in subglacial systems are limited and not contrasted with subaerial conditions at the glacier margins.

@@

\begin{table}[ht]
  \centering
  \caption{Variables, parameters, and constants }
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

@@

Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
The formulation here does not consider the englacial storage of water, and thus that changes in the hydraulic head at the top of the glacier are negligible in the evolution of the channel size. 
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_2 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n.
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the second term represents the creep closure of the channel from ice deformation.

@@

With knowledge of cross-sectional area $S$, the shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
\begin{equation}
  \label{eq:tau}
  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
\end{equation}
% 
where  the water velocity $v_g = \frac{Q_w}{S_g}$.

@@

The parameterizations are first applied to a reference test case for each glacier that assumes a semi-circular subglacial channel ($\beta=\pi$) and friction factors $f_r$ and $f_p$ are tuned so that the model reproduces reasonable velocities for both  the Greenland Ice Sheet and alpine glaciers \citep[$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$][]{werder2010b,chandler2013}.
Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial systems are presented below.

To characterize the variability in sediment discharge capacity in subglacial compared to subaerial systems further, the parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors $f_r$ and $f_p$.
Model runs with parameter combinations  are culled if their maximum subglacial water velocity exceeds $2\,$\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity exceeds $1\,$\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with some dye tracing measurements \citep[e.g.]{werder2010b,chandler2013}.
The standard deviation of water velocity, shear stress, and width-integrated shear stress across the season-long hydrographs are presented below.

@@

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig2.pdf}
    \caption{Model outputs as in response to hydrograph (a and e) from \alpine{} (a-d) and \icesheet{} (e-h). Gray lines in a and e represent the subglacial channel cross-sectional area. Below this blue lines represent model outputs of velocity, shear stress, and width-integrated shear stress from the subglacial system, while orange lines represent the subaerial system.
      Plots are shown at $15$\unit{min} interval.
    Arrows denote some peak events where peak the variable peaks in the subglacial system prior to the subaerial one.} 
    \label{fig:model_outs}
  \end{figure}
\end{center}

Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and total shear stress exhibit different seasonal evolutions and peaks of water velocity and shear stress (representing the potential for sediment mobilization) and total shear stress (representing the total sediment transport capacity; Figure~\ref{fig:model_outs}).

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig3.png}
    \caption{Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for the \alpine{} hydrograph (a-c) and \icesheet (d-f). Variables have been normalized to mean values for better comparison amongst the model runs.
      Plots are shown at $15$\unit{min} interval.} 
    \label{fig:Qw_vari}
  \end{figure}
\end{center}

For each water discharge value in the subaerial system, there is a single water velocity, shear stress and total shear stress (Figure~\ref{fig:Qw_vari}).
Conversely in the subglacial system,  water velocity, shear stress and total shear stress vary substantially for a given water discharge (Figure~\ref{fig:Qw_vari}).
For instance, over $15$ minute time periods at \alpine{}, the range of water velocities and shear stress spans from those below the mean to those almost $2$ times the mean velocity.
Substantially high values of water velocity and shear stress can occur from minimal water discharge at $2.5$\,\unit{m}$^3$\,\unit{s}$^{-1}$ to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
At \icesheet, reduced diurnal variability in the hydrograph  (Figure~\ref{fig:model_outs} e) results in smaller variations in water velocity and shear stress compared to \alpine{} (Figure~\ref{fig:Qw_vari}).
However, mean values of water velocity can occur at water discharges between roughly $50$ \,\unit{m}$^3$\,\unit{s}$^{-1}$ and $225$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, effectively spanning the observed range of water discharge.

When accounting for the evolving channel width of the subglacial conduit, total shear stress generally increases with water discharge.
Yet the total shear stress across the can vary substantially, with the highest values in total shear stress occurring at water discharge values ranging from roughly $11$ \,\unit{m}$^3$\,\unit{s}$^{-1}$, to the maximum water discharge at over $17$ \,\unit{m}$^3$\,\unit{s}$^{-1}$.
The variance is less at the \icesheet case, yet a range of total shear stress can occur at a given discharge(Figure~\ref{fig:Qw_vari} c, e).

Over the course of the season, large peaks in water velocity and shear stress, along with total shear stress, occur in response to increases in water discharge in both systems.
In the subglacial system, these peaks generally occur over the period when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
Conversely, in the subaerial system peaks in these variables occur when water discharge is at its highest value.
As a result, peaks in the subglacial variables occur prior to the peaks in the subaerial variables, representing lags in sediment mobilization and  transport capacity when sediment is transferred from one system to the other.

\begin{center}
  \begin{figure}[h]
    \includegraphics[width=0.8\linewidth]{Fig4.png}
    \caption{Variance (standard deviation) in water velocity, shear stress and total shear stress against aggregation time for the subglacial (blue) and subaerial (orange) systems from the \alpine{} (a-c) and \icesheet{} (d-f).
      Shaded areas denote the $25^{\mathrm{th}}$ and $75^{\mathrm{th}}$ quantiles of the $100$ parameter combinations for each of the subglacial and subaerial systems.
      Solid lines denote the mean values.
      Time intervals show in markers denote $15$\,\unit{min}, $1$\,\unit{hr}, $12$\,\unit{hr}, $1$\,\unit{d}, $5$\,\unit{d}, $10$\,\unit{d}, $15$\,\unit{d}, and $30$\,\unit{d} aggregation periods.
    } 
    \label{fig:multi_run}
  \end{figure}
\end{center}

Variability in water velocity and shear stress remain higher in the subglacial system compared to the subaerial one, across the range of parameter values examined for both systems.
However, in the \alpine{} system accounting for channel width makes the variability in the systems comparable (Figure~\ref{fig:multi_run}).
Yet, the variability over time periods is less than one-half day, and the variability in the subglacial system is somewhat higher. 
Conversely, in \icesheet{}, with lower water discharge variability, the variability in total shear stress remains higher in the subglacial system across the time aggregation periods.

In all cases, the variability in velocity, shear stress, and total shear stress decrease after $1-5$ day periods (Figure~\ref{fig:multi_run}).
This suggests that the effects of the variable response to hydrologic forcing in these cases are reduced after these time periods.
Yet, substantial differences persist in periods of up to $30$ days, as the variability might not converge given seasonal variations in water discharge.
Note that lags and time of peak sediment transport conditions will remain offset (Figure~\ref{fig:model_outs}), while the variability in sediment transport parameters is reduced with increasing time aggregation (Figure~\ref{fig:multi_run}).

@@

Both subglacial and subaerial parameterizations here assume that the distribution of velocity and shear stress of water across the channel bed is homogeneous \citep[Section~\ref{sect:sub_mode}~and~\ref{sect:fluv}; e.g. ]{yager2018}.
In the subaerial parameterization, this means that sediment is transported across all channel widths, as opposed to large channel widths during floods when much sediment transport occurs \citep{wolman1960}.
As a result, the  subaerial systems variability in sediment transport capacity across the range of discharge values could be underestimated in the parameterization here (Figure~\ref{fig:Qw_vari}).
Furthermore, the similar variance of total shear stress in some cases shows (Figure~\ref{fig:multi_run}) that subglacial channels contain self-organizing feedbacks that limit sediment transport capacity, as in fluvial systems \citep{phillips2016}.

In natural systems, variable mobilization and deposition of sediment may accumulate as water and sediment accumulate through the two-dimensional channel network will compound the already substantial variability in sediment mobilization parameters accounted for here in the non-dimensional parameterization.
Such variability could be especially strong in the subglacial system \citep{delaney2023}.

@@

Results here suggest that increased variability in sediment transport in subglacial systems  could be particularly pronounced on ice sheets with large catchments and reduced variability.
Furthermore, total shear stress can remain high across a range of water discharges (Figures~\ref{fig:model_outs} and \ref{fig:Qw_vari}).
In fact, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}.
Such a large quantity of sediment transported at lower water discharges may occur in part because sediment transport capacity responds to the size of the glacier conduit, in addition to water discharge (Figure~\ref{fig:model_outs} and Section~\ref{sect:sub_mode}).

This suggests that larger climatic perturbations that isolate  water discharge might be needed to isolate significant changes in sediment discharge from glacial systems compared to subaerial on, as the noise resulting from increased variability in sediment transport capacity must be overcome.
Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from  water pressure variations and sliding  \citep{iverson2012,herman2015}, which are also complex.
